### 2023-09-29 proofs/main.tex
Resolve reference issues

The specification is represented by an inference system of
three mutually recursive judgments:
positive inference $[[Γ ; Φ ⊢ v : iP]]$,
negative type inference $[[Γ ; Φ ⊢ c : iN]]$, and
application type inference $[[Γ ; Φ ⊢ iN ● args ⇒> iM ]]$.
In the premises, the inference rules also refer to 
the declarative subtyping (\cref{def:subDOne}),
type well-formedness (\cref{alg:wf}), 
and normalization (\cref{alg:alg:type-nf}) judgements.
\begin{itemize}
  \item $[[Γ ; Φ ⊢ v : iP]]$ (and symmetrically, $[[Γ ; Φ ⊢ c : iN]]$)
    means that under the type context $[[Γ]]$ and
    the variable context $[[Φ]]$, for the value $[[v]]$,
    type $[[iP]]$ is inferrable. It guarantees that 
    $[[v]]$ is well-formed in $[[Γ]]$ and $[[Φ]]$ in
    the standard sense.
  \item $[[ Γ ; Φ ⊢ iN ● args ⇒> iM ]]$ is the application type inference 
    judgement. It means that if a head of type $[[iN]]$ 
    is applied to list of values $[[args]]$, 
    then the resulting computation can be typed as $[[iM]]$.
\end{itemize}

@@

Mirroring the declarative typing, 
the algorithm is represented by an inference system of three mutually recursive
judgments:
\begin{itemize}
  \item $[[Γ ; Φ ⊨ v : iP]]$ and  $[[Γ ; Φ ⊨ c : iN]]$
    are the algorithmic versions of $[[Γ ; Φ ⊢ v : iP]]$ and $[[Γ ; Φ ⊢ c : iN]]$.
    In contrast with the declarative counterparts, they are deterministic,
    and  guarantee that the inferred type is normalized. 
  \item $[[Γ ; Φ ; Θ1 ⊨ uN ● args ⇒> uM ⫤ Θ2 ; SC]]$
    is the algorithmization of $[[Γ ; Φ ⊢ iN ● args ⇒> iM]]$.
    Notice that $[[uN]]$ contains algorithmic variables, 
    which are specified by the context $[[Θ1]]$.
    Moreover, the inferred type $[[uM]]$ is also algorithmic,
    and can have several non-equivalent instantiations. To accommodate that, 
    the algorithm also returns $[[Θ2]]$ and $[[SC]]$ specifying 
    the variables used in $[[uM]]$: $[[Θ2]]$ defines the contexts
    in which the variables must be instantiated, and $[[SC]]$
    imposes restrictions on the variables. 
\end{itemize}
As subroutines, the algorithm calls
subtyping (\cref{alg:subtyping}),
type well-formedness (\cref{alg:wf}),
constraint merge (\cref{sec:constraint-merge}),
normalization (\cref{alg:type-nf}),
and constraint singularity which will be defined later in \cref{todo}.
It also relies on basic set operations and the ability to 
deterministically choose fresh variables.

@@

\subsection{Constraint Singularity}

>>>>

The specification is represented by an inference system of
three mutually recursive judgments:
positive inference $[[Γ ; Φ ⊢ v : iP]]$,
negative type inference $[[Γ ; Φ ⊢ c : iN]]$, and
application type inference $[[Γ ; Φ ⊢ iN ● args ⇒> iM ]]$.
In the premises, the inference rules also refer to 
the declarative subtyping (\cref{def:subDOne}),
type well-formedness (\cref{alg:wf}), 
and normalization (\cref{alg:type-nf}).
\begin{itemize}
  \item $[[Γ ; Φ ⊢ v : iP]]$ (and symmetrically, $[[Γ ; Φ ⊢ c : iN]]$)
    means that under the type context $[[Γ]]$ and
    the variable context $[[Φ]]$, for the value $[[v]]$,
    type $[[iP]]$ is inferrable. It guarantees that 
    $[[v]]$ is well-formed in $[[Γ]]$ and $[[Φ]]$ in
    the standard sense.
  \item $[[ Γ ; Φ ⊢ iN ● args ⇒> iM ]]$ is the application type inference 
    judgement. It means that if a head of type $[[iN]]$ 
    is applied to list of values $[[args]]$, 
    then the resulting computation can be typed as $[[iM]]$.
\end{itemize}

@@

Mirroring the declarative typing, 
the algorithm is represented by an inference system of three mutually recursive
judgments:
\begin{itemize}
  \item $[[Γ ; Φ ⊨ v : iP]]$ and  $[[Γ ; Φ ⊨ c : iN]]$
    are the algorithmic versions of $[[Γ ; Φ ⊢ v : iP]]$ and $[[Γ ; Φ ⊢ c : iN]]$.
    In contrast with the declarative counterparts, they are deterministic,
    and  guarantee that the inferred type is normalized. 
  \item $[[Γ ; Φ ; Θ1 ⊨ uN ● args ⇒> uM ⫤ Θ2 ; SC]]$
    is the algorithmization of $[[Γ ; Φ ⊢ iN ● args ⇒> iM]]$.
    Notice that $[[uN]]$ contains algorithmic variables, 
    which are specified by the context $[[Θ1]]$.
    Moreover, the inferred type $[[uM]]$ is also algorithmic,
    and can have several non-equivalent instantiations. To accommodate that, 
    the algorithm also returns $[[Θ2]]$ and $[[SC]]$ specifying 
    the variables used in $[[uM]]$: $[[Θ2]]$ defines the contexts
    in which the variables must be instantiated, and $[[SC]]$
    imposes restrictions on the variables. 
\end{itemize}
As subroutines, the algorithm calls
subtyping (\cref{alg:subtyping}),
type well-formedness (\cref{alg:wf}),
constraint merge (\cref{sec:constraint-merge}),
normalization (\cref{alg:type-nf}),
and constraint singularity which will be defined later in 
\cref{sec:constraint-singularity}.
It also relies on basic set operations and the ability to 
deterministically choose fresh variables.

@@

\subsection{Constraint Singularity}
\label{sec:constraint-singularity}

### 2023-09-29 proofs/tex/decl-subtyping-lemmas.tex
Resolve reference issues

      Notice that since the derivation of $[[Γ1, pbs ⊢ [iPs/pas]iN' ≤ iM']]$ is
      a subderivation of the derivation of $[[Γ ⊢ iN ≤ iM]]$, its size is smaller, 
      and hence, the induction hypothesis applies
      ($[[Γ1, pbs ⊢ σ : Γ1, pbs]]$ by  \label{lemma:subst-domain-weakening})
      :
      $[[Γ2, pbs ⊢ [σ][iPs/pas]iN' ≤ [σ]iM']]$.

>>>>

      Notice that since the derivation of $[[Γ1, pbs ⊢ [iPs/pas]iN' ≤ iM']]$ is
      a subderivation of the derivation of $[[Γ ⊢ iN ≤ iM]]$, its size is smaller, 
      and hence, the induction hypothesis applies
      ($[[Γ1, pbs ⊢ σ : Γ1, pbs]]$ by \cref{lemma:subst-domain-weakening})
      :
      $[[Γ2, pbs ⊢ [σ][iPs/pas]iN' ≤ [σ]iM']]$.

### 2023-09-29 proofs/tex/equiv-lemmas.tex
Resolve reference issues

  \label{lemma:poly-types-equivalence}
      Notice that the application of bijection $[[μ]]$ to $[[iM']]$ does
      not change its size (which is less than the size of $[[iM]]$), hence the induction hypothesis applies.
      This way, $[[ [Sub μ]iM' ≈ iN']]$ (and by symmetry, $[[iN' ≈ [Sub μ]iM']]$) holds by induction. 
      Then we apply \ruleref{\ottdruleEOneForallLabel} to get $[[∀pas.iN' ≈ ∀pbs.iM']]$, i.e. $[[iN ≈ iM]]$.
    \end{caseof}

@@

\begin{lemma}[Substitution preserves declarative equivalence]
  \label{corollary:subst-pres-equiv}
  Suppose that $[[σ]]$ is a substitution. Then
    \begin{itemize}
      \item[$+$] $[[iP ≈ iQ]]$ implies $[[ [σ]iP ≈ [σ]iQ ]]$
      \item[$-$] $[[iN ≈ iM]]$ implies $[[ [σ]iN ≈ [σ]iM ]]$
  \end{itemize}
\end{lemma}
\begin{proof}

>>>>

      Notice that the application of bijection $[[μ]]$ to $[[iM']]$ does
      not change its size (which is less than the size of $[[iM]]$), hence the induction hypothesis applies.
      This way, $[[ [Sub μ]iM' ≈ iN']]$ (and by symmetry, $[[iN' ≈ [Sub μ]iM']]$) holds by induction. 
      Then we apply \ruleref{\ottdruleEOneForallLabel} to get $[[∀pas.iN' ≈ ∀pbs.iM']]$, i.e. $[[iN ≈ iM]]$.
    \end{caseof}

@@

\begin{lemma}[Substitution preserves declarative equivalence]
  \label{corollary:subst-pres-decl-equiv}
  Suppose that $[[σ]]$ is a substitution. Then
    \begin{itemize}
      \item[$+$] $[[iP ≈ iQ]]$ implies $[[ [σ]iP ≈ [σ]iQ ]]$
      \item[$-$] $[[iN ≈ iM]]$ implies $[[ [σ]iN ≈ [σ]iM ]]$
  \end{itemize}
\end{lemma}
\begin{proof}

### 2023-09-29 proofs/tex/upgrade-lemmas.tex
Resolve reference issues

    Let us apply $[[pnas/pnbs]]$---the inverse of the substitution $[[ pnbs/pnas ]]$ to 
    both sides of $[[Δ, pnbs, pncs ⊢ iQ ≥ [pnbs/pnas]iP]]$ and 
    by \cref{lemma:subst-pres-subt} 
    (since $[[pnbs/pnas]]$ can be specified as 
    $[[Δ,pnbs,pncs ⊢ pnbs/pnas : Δ, pnas, pncs]]$ by \label{lemma:subst-domain-weakening})
    obtain $[[Δ, pnas, pncs ⊢ [pnas/pnbs]iQ ≥ iP]]$.
    Notice that $[[Δ ⊢ iQ]]$ implies that $[[fv iQ ∩ {pnbs} = ∅]]$, 
    then by \cref{corollary:subst-disj}, $[[ [pnas/pnbs]iQ = iQ]]$,
     and thus $[[Δ, pnas, pncs ⊢ iQ ≥ iP]]$.
    By context strengthening, $[[Δ, pnas ⊢ iQ ≥ iP]]$.
\end{proof}

>>>>

    Let us apply $[[pnas/pnbs]]$---the inverse of the substitution $[[ pnbs/pnas ]]$ to 
    both sides of $[[Δ, pnbs, pncs ⊢ iQ ≥ [pnbs/pnas]iP]]$ and 
    by \cref{lemma:subst-pres-subt} 
    (since $[[pnbs/pnas]]$ can be specified as 
    $[[Δ,pnbs,pncs ⊢ pnbs/pnas : Δ, pnas, pncs]]$ by \cref{lemma:subst-domain-weakening})
    obtain $[[Δ, pnas, pncs ⊢ [pnas/pnbs]iQ ≥ iP]]$.
    Notice that $[[Δ ⊢ iQ]]$ implies that $[[fv iQ ∩ {pnbs} = ∅]]$, 
    then by \cref{corollary:subst-disj}, $[[ [pnas/pnbs]iQ = iQ]]$,
     and thus $[[Δ, pnas, pncs ⊢ iQ ≥ iP]]$.
    By context strengthening, $[[Δ, pnas ⊢ iQ ≥ iP]]$.
\end{proof}
